Molecular dynamics simulation of synchronization in driven particles

### 2020-12-23 manuscript/manuscript_ajp_2020.tex
updated text and references

\author{Tiare Guerrero}
\email{guer9330@pacificu.edu} % optional
%\altaffiliation[permanent address: ]{101 Main Street, 
%  Anytown, USA} % optional second address
% If there were a second author at the same address, we would put another 
% \author{} statement here.  Don't combine multiple authors in a single
% \author statement.
\affiliation{Department of Physics, Pacific University, Forest Grove, OR 97116}
% Please provide a full mailing address here.

@@

\begin{abstract}
  %What problem did you study and why is it important?
  We discuss a
  numerical model
  of particles
  confined to a narrow channel 
  driven
  across a washboard potential energy landscape.
  The model 
  exhibits
  synchronization effects that provide insight
  to the behavior of 
  many experimental systems of
  interest to condensed matter physics 
  including 
  magnetically driven colloidal particles
  confined by light-fields, 
  voltage driven superconducting vortices
  confined in a Josephson junction,
  and ac and dc driven
  charge and spin density waves.
  We present the basics
  of the molecular dynamics simulations
  for particles moving through
  a viscous liquid.
  %so we model the dynamics of each particle
  %with 
  %overdamped equations of motion.
  We demonstrate how to visualize
  the system
  and measure hallmarks of synchronization, 
  Shapiro steps,
  that exhibit non-Ohmic step-like behavior 
  in the 
  the particle velocity as a function of applied driving force.
  We include sample code
  and exercises for students
  %that include
  with 
  opportunties
  to reproduce our results and propose
  new numerical experiments.
  With only a few particles in two-dimensions,
  the simulation runs quickly,
  making this an appropriate model for undergraduates to explore.

@@

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
in which individual oscillators adjust rhythm due
to an external stimuli \cite{Pikovsky2003}.
%Christiaan Huygens 
%[has been adapted into many useful technologies...]
Synchronization was first reported in the scientific literature
when in 1665 
Huygens' experiments on
the 
motions of two pendula on wall-mounted clocks 
demonstrated the tendency of 
two periodic oscillators in close
proximity to swing in time, i.e. to synchronize,
due to interactions through the wall \cite{Bennett2002}.
Synchronous behaviors are observed
in many everyday systems
such as the
flickering candle flames coupled by temperature fluctuations \cite{Okamoto2016} 
and metronomes coupled through the the supporting table \cite{Jia2015}.
Biological systems benefit from cooperative
synchronzation such as when 
birds coordinate
flapping of wings in order to optimize energy use during flight \cite{Portugal2014},
frogs communicate by croaking patterns dictated by location \cite{Aihara2014},
and humans synchronize clapping in time with music \cite{Tranchant2016}.
At a cellular level, 
neurons simultaneously fire in cardiac muscle \cite{MartinHall1999}
and brain tissue \cite{Singer1999}.

Synchronization is also referred to as 
phase-locking or mode-locking.
%phase locking
Dynamical systems %with multiple frequencies
exhibit phase-locking 
when oscillators with different frequencies couple
so the frequency ratio or mode 
is an integer number \cite{Bak1986} .
This can be demonstrated
with Lissajous figures, i.e. parametric graphs of two periodic functions
which make looped patterns determined by the mode, % frequency ratio,
first reported in 1857 \cite{Lissajous1857}
and easily generated with a computer code 
or oscilloscope \cite{Tong1997}. 
%with a frequency generator,
%where two combined frequencies at perpendicular orientations

Synchronization is often achieved via external forcing.
Examples of synchronization via external forcing include
pushing a child on a swing or
using a pacemaker (an electrical generator)
that pulses to regulate a heart beat.
This kind of 
dynamical mode locking
is often observed in quantum electronic
devices such as Josephson junctions \cite{Josephson1962,Josephson1965}
i.e. phase-locked current loops \cite{} 
which exhibit phase locking via
stepped regions in current-voltage (I-V) 
instead of ohmic relationships.
Also known as Shapiro steps \cite{Shapiro1963},
these steps have been observed broadly in 
velocity-force curves and voltage-current curves
as indicators of phase locking \cite{Golubov2004}.

Recent work with atomic systems such as Bose-Einstein condensates \cite{Zhang2020}

Synchronization has many important
technological applications
and is often studied with simplified computational models.
%mini review of synchronization in electronics
%careful - the following is all simulations
Recent studies demonstrate
that coupled electronic oscillators can be
programmed for use as microcontrollers for robotic locomotion \cite{Dutta2019}.

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we focus numerical studies 
of the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape. 
We chose this model for its
relevance to condensed matter systems

Studies modeling synchronized behaviors
can be performed with
coupled oscillators []
An alternative method is to
study the dynamics of a single particle
confined in a periodic 
potential landscape subject
to with an applied driving force with constant or direct (DC)
and alterating (AC)
components.

Relationship between substrate period and intrinsic velocity
with phase locking caused by AC force modulates drive
where the particle velocity is constant even though the applied
DC force is increased.

Recent experiments of particles confined
in optical traps
subject to external driving forces 
have
been used to examine the microscopic dynamics
of mode-locking in colloids \cite{juniper2015} and ions \cite{Bushev2013}.
%experimental - colloids specifically
Colloidal particles trapped in light fields are %have proven
a particularly useful medium for studying complex dynamical behaviors.  %cite 
The relatively large size of the colloids (micrometers)
makes colloids easy to control and image,
and the interaction forces between colloids can be modified
by tuning the chemistry of the suspending fluid or surface ligands
of the colloidal particles \cite{Grier2003}.
This ease in control of colloids has lead to a rich array of 
experimental results  
considered models
for experimental systems relatively hard to access and visualize,
such as cold atoms or electron gases. %ELABORATE AND CITE

@@

%external forces
The
dynamics of particles subject to 
an applied external force apply to many physical systems.
For instance the flow of charges in a conductor,
or cells responding to a chemical gradient.
The external force 
increases the diversity of dynamical behaviors,
and can cause particles to flow in
a variety of non-linear complex patterns
that may be 
synchronized.
Disordered chaotic dynamics are also possible,
where irregular, unpredictable time evolution of
nonlinear systems and occurs in mechanical oscillators \cite{chaos}.

@@

@@

%describe types of physical systems - particle types
In this work,
we perform 
numerical simulations of confined, driven particles
to model a variety of physical phenomena.
%----------------------------------------------------------------------------
%paper outline
In the following paper we describe
our molecular dynamics model in Section~\ref{sec:MD}.
%including various  and confining environments.
We include code to simulate
and visualize the dynamics in this section
and supplementary material.
In Section~\ref{sec:results} we summarize
our results,
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one} and
multiple interacting particles
in Section~\ref{sec:sync},
including stationary propagation of high density kinks
in Sec.~\ref{sec:kink}.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of velocity versus position.
In Section~\ref{sec:quasiperiod},
we show how an aperiodic landscape modifies the particle dynamics.
Finally we explore the transition to chaotic dynamics in 
Section~\ref{sec:chaos}
and conclude 
in Section~\ref{sec:conclusion}. 
In each section
we suggest exercises for interested students,
as summarized in Section~\ref{sec:problems}.

@@

We drive a single particle across a periodic landscape 
along the y-direction 
\begin{equation}
  \label{eq:landscape}
  V_l(y) = V_{0y} \cos{(N_p \pi y / L)} 
\end{equation}
with $N_p=20$ troughs in the landscape.
%and $V_{0y}=0.2$.
This is illustrated in Fig.~\ref{fig:1}(a) where
the red (blue) regions show local maxima (minima).
The code for generating
a two dimensional colored plot
of the potential landscape
is calculated by evaluating
the analytic function in Eq.~\ref{eq:landscape}
for a grid of values $(x_n,y_n)$.
%is included in the supplementary material,
The numerical implementation of the landscape 
is calculated with
\begin{equation}
  \label{eq:landscape}
  F_y(y) = F_{0y} \sin{(N_p \pi y / L)} 
\end{equation}
where $F_{0y} = 0.1$.
We drive the particle with
$F_{AC}=0.2$, $F_{DC}=0.01$, and $f=0.01$ cycles per time unit.
When the superposition of $F_{AC}$ and $F_{DC}$ is large enough
to overcome the barrier height of the landscape troughs,
the particle hops between troughs in the energy landscape.
In Fig.~\ref{fig:0} we
show the relationship between the applied force and particle
position.
In Fig.~\ref{fig:0}(a)
we plot the applied driving force
as a function of time
and in Fig.~\ref{fig:0}(b) 
we show the position of the particle
as a function of time.
The initial position is near $y=0$
and the initial driving force is minimum
$F_D(t=0) = F_{DC}$.
Over the time period $T/2 = 50$ %time units
the particle moves
in the positive y-direction
while $F_{D}(t) > 0 $
through three
substrate troughs, reaching a maximum of $y/\lambda = 3$.
When $F_{D}(t) < 0 $
the particle moves in the negative y-direction
reaching a minimum position
of $y/\lambda = 1$
over the period $T = 1/f = 100$.
The average velocity $\bar{v}_y$
is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
In Fig. ~\ref{fig:0}
the displacement is a single wavelength
of the substrate $y(t_0+T) - y(t_0) = \lambda$.
Thus the average velocity 
is $\bar{v}_y = \lambda f$, where $T = 1/f$.

@@

Our simulations reproduce results presented in 
Juniper {\it et al.} \cite{juniper2015}
which demonstrated
mode locking in
experiments of 
driven colloids on a
optical periodic landscape.

@@

@@

@@

    \bibitem{juniper2015} M. P. N. Juniper, A. V. Straube, R. Besseling, D. G. A. L. Aarts, and R. P. A. Dullens, Microscopic dynamics of synchronization in driven colloids. Nat. Commun. 6, 7187 (2015).

    \bibitem{Bushev2013} P. Bushev, G. Hétet, L. Slodička, D. Rotter, M. A. Wilson, F. Schmidt-Kaler, J. Eschner, and R. Blatt, Shot-Noise-Limited Monitoring and Phase Locking of the Motion of a Single Trapped Ion, Phys. Rev. Lett. {\bf 110}, 133602 (2013).

>>>>

\author{Tiare Guerrero}
\email{guer9330@pacificu.edu} % optional
%\altaffiliation[permanent address: ]{101 Main Street, 
%  Anytown, USA} % optional second address
% If there were a second author at the same address, we would put another 
% \author{} statement here.  Don't combine multiple authors in a single
% \author statement.
%\affiliation{Department of Physics, Pacific University, Forest Grove, OR 97116}
% Please provide a full mailing address here.

@@

\begin{abstract}
  %What problem did you study and why is it important?
  We discuss a
  numerical model
  of particles
  confined to a narrow channel 
  driven
  across a washboard potential energy landscape.
  The model 
  exhibits
  synchronization effects that provide insight
  to the behavior of 
  experimental systems 
  %interest to condensed matter physics 
  including 
  magnetically driven colloidal particles
  confined by light-fields.
  %and
  %atoms on surfaces.
  %voltage driven superconducting vortices
  %confined in a Josephson junction,
  %and ac and dc driven
  %charge and spin density waves.
  We present the basics
  of the molecular dynamics simulations
  for particles moving through
  a viscous liquid
  applied to single and multiple particle systems.
  %so we model the dynamics of each particle
  %with 
  %overdamped equations of motion.
  We demonstrate how to 
  visualize
  the system with
  animations of the particle motion 
  and detect synchronization using
  measurements of
  the particle velocity as a function of applied driving force.
  We include sample code
  and exercises for students
  %that include
  with 
  opportunties
  to reproduce our results and propose
  new numerical experiments.
  With only a few particles in two-dimensions,
  the simulation runs quickly,
  making this an appropriate model for undergraduates to explore.

@@

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
in which individual oscillators adjust frequency due
to interaction or an external stimulus \cite{Pikovsky2003}.
Synchronous behaviors where oscillators move in-phase (i.e. in time)
are observed
in many everyday systems
such as the
flickering candle flames coupled by temperature fluctuations \cite{Okamoto2016} 
and metronomes coupled through the the supporting table \cite{Jia2015}.
Biological systems benefit from cooperative
synchronzation such as when 
birds coordinate
flapping of wings in order to optimize energy use during flight \cite{Portugal2014},
frogs communicate by croaking patterns modulated by location \cite{Aihara2014},
and humans synchronize clapping in time with music \cite{Tranchant2016}.
At a cellular level, 
neurons simultaneously fire in cardiac muscle \cite{MartinHall1999}
and brain tissue \cite{Singer1999}.

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
A particular form of 
synchronization is phase-locking,
which first appeared in the scientific literature
with 
Huygens' 1665 experiments on
the 
motions of two pendula on wall-mounted clocks.
Huygens demonstrated the tendency of 
two periodic oscillators in close
proximity to swing in time %, i.e. to synchronize,
due to interactions through the wall 
no matter the original phase of the clocks \cite{Bennett2002}.
%phase locking
Dynamical systems %with multiple frequencies
exhibit phase-locking 
when oscillators with different frequencies couple
so the frequency ratio or mode 
is an integer number \cite{Bak1986} .
This can be demonstrated
with Lissajous figures, i.e. parametric graphs of two periodic functions
which make looped patterns determined by the mode, % frequency ratio,
first reported in 1857 \cite{Lissajous1857}
and easily generated with a computer code 
or oscilloscope \cite{Tong1997}. 
%with a frequency generator,
%where two combined frequencies at perpendicular orientations

Synchronization is often achieved via external forcing
such as when
%single particle oscillators
pushing a child on a swing at the high point on her path, or
using pacemaker (an electrical generator)
pulses to regulate a heart beat.
%multiple particle oscillators
[Fireflies and pulsed LED].

Dynamical mode locking 
has many important
technological applications
and is often studied with simplified computational models.
These studies can be performed
with coupled oscillators \cite{Kuramoto193?}
or with a single particle driven across a periodic potential landscape \cite{}.
%Single particle models avoid many-body effects
%while offering insight.

%Oscillators ordered into chains or lattices.
Synchronization 
is %often
observed
in quantum electronic
devices such as Josephson junctions \cite{Josephson1962,Josephson1965}
i.e. two superconductors sandwiching an insulating layer.
When subject to an external voltage,
Cooper pairs in the superconducting materials
tunnel through the insulating layer.
Phase-locking is observed as 
%phase-locked current loops \cite{} 
stepped regions in current-voltage (I-V) relationship in these devices,
where a constant current results from a range of voltages
due to the quantized nature of the tunneling probability [NOT QUITE RIGHT...].
%instead of ohmic relationships.
Known as Shapiro steps \cite{Shapiro1963},
these indicators of phase locking 
have been observed broadly in a variety of systems
either in 
velocity-force curves and voltage-current curves \cite{Golubov2004}.
Shapiro steps
are hallmarks of
non-Ohmic behavior in voltage-current curves,
where these steps are generated
by the tendency of the response of a system
to be flat across a range of applied driving forces or voltages,
which can be probed with applied AC forces.

%  atoms on surfaces.
%  voltage driven superconducting vortices
%  confined in a Josephson junction,
%  and ac and dc driven
%  charge and spin density waves.
%
%
%mini review of synchronization in electronics
%careful - the following is all simulations
%Recent studies demonstrate
%that coupled electronic oscillators can be
%programmed for use as microcontrollers for robotic locomotion \cite{Dutta2019}.

Recent experiments of particles confined
in optical traps
subject to external driving forces 
have
been used to examine the microscopic dynamics
of mode-locking in colloids interacting via magnetic dipole
or electrostatic interactions
in experiment \cite{Juniper2015, Lutz2004} and
simulation \cite{Herrera-Velarde2007, Herrera-Velarde2008}.
%experimental - colloids specifically
Colloidal particles trapped in light fields are %have proven
a particularly useful medium for studying complex dynamical behaviors.  %cite
Colloid particles are micron-sized,
a relatively large size to control and image.
The interaction forces between colloids can be modified
by tuning the chemistry of the suspending fluid or surface ligands
of the colloidal particles \cite{Grier2003}.
This ease in control of colloids has lead to a rich array of 
experimental results  
considered models
for experimental systems relatively hard to access and visualize,
such as cold atoms or electron gases. %ELABORATE AND CITE
%Recent work with atomic systems such as Bose-Einstein condensates \cite{Zhang2020}

@@

%external forces
The
dynamics of particles subject to 
an applied external force apply to many physical systems.
For instance the flow of charges in a conductor,
or cells responding to a chemical gradient.
The external force 
increases the diversity of dynamical behaviors,
and can cause particles to flow in
a variety of non-linear complex patterns
that may be 
synchronized.
Disordered chaotic dynamics are also possible,
where irregular, unpredictable time evolution of
nonlinear systems occurs in mechanical oscillators \cite{chaos}.

@@

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we focus numerical studies 
of the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape. 
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%We examine
The period of the substrate
can be used to control the
intrinsic velocity of the driven particle.

@@

%describe types of physical systems - particle types
In this work,
we perform 
numerical simulations of confined, driven particles
to model a variety of physical phenomena.
%----------------------------------------------------------------------------
%paper outline
In the following paper we describe
our molecular dynamics model in Section~\ref{sec:MD}.
%including various  and confining environments.
We include code to simulate
and visualize the dynamics in this section
and supplementary material.
In Section~\ref{sec:results} we summarize
our results,
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one} and
multiple interacting particles
in Section~\ref{sec:sync},
including stationary propagation of high density kinks
in Sec.~\ref{sec:kink}.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWNING COUPLING, OR QUASIPERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
In Section~\ref{sec:quasiperiod},
we show how an aperiodic landscape modifies the particle dynamics.
Finally we explore the transition to chaotic dynamics in 
Section~\ref{sec:chaos}
and conclude 
in Section~\ref{sec:conclusion}. 
In each section
we suggest exercises for interested students,
as summarized in Section~\ref{sec:problems}.

@@

We drive a single particle across a periodic landscape 
along the y-direction 
\begin{equation}
  \label{eq:landscape}
  V_l(y) = V_{0y} \cos{(N_p \pi y / L)} 
\end{equation}
%v is the standard I note for substrates
with $N_p=20$ troughs in the landscape.
%and $V_{0y}=0.2$.
This is illustrated in Fig.~\ref{fig:1}(a) where
the red (blue) regions show local maxima (minima).
The code for generating
a two dimensional colored plot
of the potential landscape
is calculated by evaluating
the analytic function in Eq.~\ref{eq:landscape}
for a grid of values $(x_n,y_n)$.
%is included in the supplementary material,
The numerical implementation of the landscape 
is calculated with
\begin{equation}
  \label{eq:landscape}
  F_y(y) = F_{0y} \sin{(N_p \pi y / L)} 
\end{equation}
where $F_{0y} = 0.1$.
We drive the particle with
$F_{AC}=0.2$, $F_{DC}=0.01$, and $f=0.01$ cycles per time unit.
When the superposition of $F_{AC}$ and $F_{DC}$ is large enough
to overcome the barrier height of the landscape troughs,
the particle hops between troughs in the energy landscape.
In Fig.~\ref{fig:0} we
show the relationship between the applied force and particle
position.
In Fig.~\ref{fig:0}(a)
we plot the applied driving force
as a function of time
and in Fig.~\ref{fig:0}(b) 
we show the position of the particle
as a function of time.
The initial position is near $y=0$
and the initial driving force is minimum
$F_D(t=0) = F_{DC}$.
Over the time period $T/2 = 50$ %time units
the particle moves
in the positive y-direction
while $F_{D}(t) > 0 $
through three
substrate troughs, reaching a maximum of $y/\lambda = 3$.
When $F_{D}(t) < 0 $
the particle moves in the negative y-direction
reaching a minimum position
of $y/\lambda = 1$
over the period $T = 1/f = 100$.
The average velocity $\bar{v}_y$
is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
In Fig. ~\ref{fig:0}
the displacement is a single wavelength
of the substrate $y(t_0+T) - y(t_0) = \lambda$.
Thus the average velocity 
is $\bar{v}_y = \lambda f$, where $T = 1/f$.

@@

Our simulations reproduce results presented in 
Juniper {\it et al.} \cite{Juniper2015}
which demonstrated
mode locking in
experiments of 
driven colloids on a
optical periodic landscape.

@@

\begin{enumerate}
\item Kuramoto model = overdamped interacting rotors/oscillators are used to model coupled oscillator systems - no substrate or external drive so far as I can see - this seems baked into the oscillator properties/interactions
%D. Métivier, L. Wetzel, and S. Gupta
%Onset of synchronization in networks of second-order Kuramoto oscillators with delayed coupling: Exact results and application to phase-locked loops
%PHYSICAL REVIEW RESEARCH 2, 023183 (2020)

\item Brownian motion without fluid dynamics using  Ermak's algorithm

\item Frenkel-Kontorova and competing length scales for low temperature systems where thermal fluctuations are neglible

  \item The Fokker-Planck Equation

\end{enumerate}

@@

%  Coupling between two singing wineglasses
%Tal Arane, Ana K. R. Musalem and Moti Fridman
%American Journal of Physics 77, 1066 (2009); https://doi.org/10.1119/1.3119175

%Synchronization of a thermoacoustic oscillator by an external sound source
%G. Penelet and T. Biwa
%American Journal of Physics 81, 290 (2013); https://doi.org/10.1119/1.4776189

@@

    \bibitem{Juniper2015} M. P. N. Juniper, A. V. Straube, R. Besseling, D. G. A. L. Aarts, and R. P. A. Dullens, Microscopic dynamics of synchronization in driven colloids. Nat. Commun. 6, 7187 (2015).

    \bibitem{Herrera-Velarde2008} S. Herrera-Velarde and R. Castañeda-Priego, Superparamagnetic colloids confined in narrow corrugated substrates, PHYSICAL REVIEW E {\bf 77}, 041407 (2008).

    \bibitem{Herrera-Velarde2007} S. Herrera-Velarde and R. Castañeda-Priego, J. Phys.: Condens. Matter 19, 226215 (2007).

%      Lutz et al. 关11,12兴 created 1D circular channels by means of scanning optical tweezers in order to avoid the presence of lateral confinement walls
    \bibitem{Lutz2004} C. Lutz, M. Kollmann, and C. Bechinger, Phys. Rev. Lett. {\bf 93}, 026001 (2004); C. Lutz, M. Kollmann, C. Bechinger, and P. Leiderer, J. Phys.: Condens. Matter {\bf 16}, S4075 (2004).
